---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading required packages
```{r, loadingpackages, results='hide'}
library(tidyverse)
library(xtable)
```

## Loading and preprocessing the data

```{r, setWD, echo=FALSE}
setwd("C:/Users/BusinessTim/Desktop/desktop files/Coursera/ReProReseach/Project1/RepData_PeerAssessment1/")

```

Loading the data
```{r, loadingdata}
activity.data<-read_csv("activity.csv")
```

## What is mean total number of steps taken per day?

Calculate total steps per day

```{r, results='hide'}
total.daily.steps<-activity.data %>%
  group_by(date)%>%
  summarize(sum.daily.steps=sum(steps)) %>%
  drop_na()
```

Histogram of the total number of steps taken each day
```{r}
ggplot(data=total.daily.steps, aes(x=sum.daily.steps))+
  geom_histogram(fill='blue', color='black')+
    xlab('Total daily steps')+
    ylab('Count')
```

Mean and median of the total number of steps taken per day

```{r, results='asis'}
daily.steps.summary.stats<-total.daily.steps %>%
  ungroup() %>%
  summarize(mean.daily.step=mean(sum.daily.steps, na.rm=TRUE), 
            median.daily.steps=median(sum.daily.steps, na.rm=TRUE))%>%
  rename(Mean_Daily_Steps=mean.daily.step,
         Median_Daily_Steps=median.daily.steps)

ds.xt<-xtable(daily.steps.summary.stats)

print(ds.xt, type='html')
```


## What is the average daily activity pattern?

Average number of steps taken per 5 minute interval:

```{r, results='hide'}
mean.interval.steps<-activity.data %>%
  group_by(interval)%>%
  summarize(mean.interval.steps=sum(steps, na.rm=TRUE)) 
```

```{r}
(ggplot(data=mean.interval.steps, aes(x=interval, y=mean.interval.steps))+
    geom_line()+
    ylab('Mean steps per 5-minute interval')+
    xlab('Interval'))
```

Interval with the maximum mean steps:

```{r, results='asis'}
max.steps<-mean.interval.steps %>%
  filter(mean.interval.steps==max(mean.interval.steps)) %>%
  dplyr::select(interval)

print(xtable(max.steps), type='html')
```


## Imputing missing values

Rows with missing data:

```{r}
        missing.data.rows<-activity.data %>%
          mutate(missing=is.na(steps)) %>%
          filter(missing==TRUE) %>%
          nrow()
        
        print(missing.data.rows)
```



Histogram of imputed total daily steps
```{r}
imputed.data<-data.frame()

interval.list<-unique(activity.data$interval)

for (i in interval.list){
  
      to.impute<-activity.data %>%
        filter(interval==i) 
      
      to.impute$steps[is.na(to.impute$steps)]<-mean(to.impute$steps, na.rm=TRUE)
      
      imputed.data<-bind_rows(imputed.data, to.impute)

}

imputed.activity.data<-imputed.data %>%
  group_by(date) %>%
  summarize(sum.daily.steps=sum(steps))

ggplot(data=imputed.activity.data, aes(x=sum.daily.steps))+
    geom_histogram(fill='blue', color='black')+
    xlab('Total daily steps')+
    ylab('Count')
```

Imputed mean and median daily steps

```{r, results='asis'}
daily.steps.summary.stats<-imputed.activity.data %>%
    ungroup() %>%
    summarize(mean.daily.step=mean(sum.daily.steps, na.rm=TRUE), 
              median.daily.steps=median(sum.daily.steps, na.rm=TRUE))

ds.xt.im<-xtable::xtable(daily.steps.summary.stats)

print(ds.xt.im, type='html')
```

The imputed mean is unaffected, but the median is now slightly higher


## Are there differences in activity patterns between weekdays and weekends?


```{r, results='hide'}
week.part<-imputed.data %>%
  mutate(day.type=as.factor(ifelse(weekdays(date)=='Sunday' | 
                                     weekdays(date)=='Saturday', 'weekend', 'weekday'))) %>%
  group_by(day.type, interval) %>%
  mutate(mean.steps=mean(steps))
```

Mean number of steps in a given 5-minute interval by weekday-day or weekend-day

```{r}
(ggplot(data=week.part, aes(x=interval, y=mean.steps))+
  geom_line()+
  facet_grid(day.type~.))
```



















